#Reproducible Research - Project Assignment 1  
____________________________________________  
##Introduction
This document shows the  use of RMarkdown for generating documentation embededed with R code

#### Step 1:  Read the input file activity.csv, change the date variable to Date type and add new column weekday name  

``` {r echo =TRUE }
activity <- read.csv("./data/activity.csv", header = TRUE, stringsAsFactors = FALSE)
activity$date <- as.Date(activity$date, "%Y-%m-%d")
activity$weekday <- weekdays(activity$date)
str(activity)
head(activity)
```

####Step 2:  Sum up steps taken for each day and plot a historgram

``` {r echo =TRUE}  

steps_per_day <- with(activity, aggregate(steps, by =list(date),FUN = sum,  na.rm=TRUE))
names(steps_per_day)= c("date" , "steps"  )
plot(steps_per_day$steps, type = "h", lwd = 2,main = " Steps Per Day", xlab = "Days", ylab = "Steps" )

```


####Step 3:  Calculate  the median and mean of steps taken each day

``` {r echo=TRUE}
mean_steps_perday   <- mean(activity$steps,na.rm = TRUE)
median_steps_perday <- median(activity$steps,na.rm = TRUE)
mean_steps_perday
median_steps_perday

```

####Step 4:  Calucate average number of steps taken per 5min-interval for all days and plot time series(line) graph  

``` {r echo=TRUE}  
meansteps_per_interval <- with(activity, aggregate(steps, by =list(interval),FUN = mean,  na.rm=TRUE))
names(meansteps_per_interval)= c("interval" , "meansteps"  )
with(meansteps_per_interval, plot(interval, meansteps, type = "l", lwd = 2,main = " Mean Steps per Interval", xlab = "Interval", ylab = "MeanSteps" ))

```  

####step 5: Find out interval that has highest average steps of 5min-interval across all days  

``` {r echo=TRUE}  
meansteps_per_interval$interval[which.max(meansteps_per_interval$meansteps)]

```   

####Step 6: Imputing missing values of steps vairable   
            Find the total number of missing values for steps variable   

``` {r echo = TRUE}
sum(!complete.cases(activity))

```   

            Create newdatset from activity dataset and replace the null steps variable with interval mean. The updated values of missing steps variable is rounded to nearest integer 
            Plot the histogram with newly created datset for steps taken daily
            Calculate the new mean and median valus for the newly created dataset
            
```  {r echo = TRUE}   
updated_activity<- transform(activity, steps = ifelse(is.na(activity$steps),round(meansteps_per_interval$meansteps[match(activity$interval,meansteps_per_interval$interval)],digits = 0),activity$steps ))  

head(updated_activity,20)

updated_steps_per_day <- with(updated_activity, aggregate(steps, by =list(date),FUN = sum,  na.rm=TRUE))
names(updated_steps_per_day)= c("date" , "steps"  )


plot(steps_per_day$steps, type = "h", col = "blue", lwd = 2,main = " Steps Per Day", xlab = "Days", ylab = "Steps" )

lines(updated_steps_per_day$steps, type = "h", col = "red", lwd = 2,main = " Steps Per Day", xlab = "Days", ylab = "Steps" )

legend("topright", c("Updated", "Original"), col=c("red", "blue"), lwd=10)

updated_mean_steps_perday   <- mean(updated_activity$steps,na.rm = TRUE)
updated_median_steps_perday <- median(updated_activity$steps,na.rm = TRUE)
updated_mean_steps_perday
updated_median_steps_perday

```   
               Differences in mean, median and totla steps  between newly createded dataset and the orignal one  
``` {r echo = TRUE}
diff_mean <- updated_mean_steps_perday - mean_steps_perday
diff_median<- updated_median_steps_perday - median_steps_perday
diff_totalsteps <- sum(updated_activity$steps,na.rm = TRUE) - sum(activity$steps, na.rm = TRUE)

diff_mean
diff_median
diff_totalsteps

```
####Step 7: Differeces in activity pattern in weekdays and weekends   
            Create datetype filed fillit it Weekday or Weekend based on Weekday  
            Load ggplot2 package and plot the graphs for Avg.Daily Steps for Weekday and Weekend
``` {r echo = TRUE}

 updated_activity$datetype[(updated_activity$weekday=="Sunday"|updated_activity$weekday=="Saturday")]<-"Weekend"

updated_activity$datetype[(updated_activity$weekday!="Sunday"&updated_activity$weekday!="Saturday")]<-"Weekday"
updated_activity_by_datetype <-aggregate(steps~interval + datetype,updated_activity, mean, na.rm = TRUE)

updated_activity_by_datetype$steps <- round(updated_activity_by_datetype$steps, digits = 2)

library(ggplot2) 
g<- ggplot(updated_activity_by_datetype, aes(x=interval, y=steps, color = datetype))+geom_line()
g<- g+labs(title="Avg.Daily Steps by Datetype", x="Interval", y= 'Avg.no. of Steps') + facet_wrap(~datetype, nrow=1,  ncol = 2)
plot(g)

```

##     End of the Script   ####



